### 2008-12-18 talk/ecoop2009/jitgen.tex
add a tlc-based example for partial evaluation.

\cfbolz{I would propose to use either TLC as an example here, or something that
looks at least like an interpreter loop}

Example::
\begin{lstlisting}[language=Python]
  def f(x, y):    
    x2 = x * x    
    y2 = y * y    
    return x2 + y2

**case x=3** ::

  def f_3(y):    
    y2 = y * y   
    return 9 + y2

**case x=10** ::

  def f_10(y):    
    y2 = y * y   
    return 100 + y2
\end{lstlisting}

A shortcoming of PE is that in many cases not much can be really assumed
constant at compile-time, and this leads to poor results.  Effective dynamic
compilation requires feedback of runtime information into compile-time; for a
dynamic language, types are a primary example.

Partial evaluation (PE) comes in two flavors:

>>>>

\begin{figure}[h]
\label{fig:tlc-main}
\begin{center}
\input{tlc-simplified.py}
\caption{The main loop of the TLC interpreter}
\end{center}
\end{figure}

Let us look at an example. Figure \ref{fig:tlc-main} shows parts of the
main-loop of the TLC interpreter (in slightly simplified form). The static
arguments of this functions would typically be the \lstinline{code} argument
(containing the bytecode as a string), the \lstinline{pc} argument (containing
the initial program counter) and the \lstinline{pool} argument (containing the
constant pool). The \lstinline{args} argument is dynamic since it contains the
user-input of the interpreted function. Since the \lstinline{while} loop and the
contained conditions depend only on static arguments it will typically be
unrolled by a partial evaluator.

When the function is partially evaluated with respect to the TLC example
function shown in Figure XXX (which computes the absolute value of a number),
the residual code would look like in Figure \ref{fig:tlc-folded}. This version
is already a great improvement over pure interpretation, all the bytecode
dispatch overhead has been removed. However, the function as shown is still
rather inefficient, due to lots of superfluous stack handling and also due to
some indirect calls for implementing comparison (\lstinline{lt}) and
subtraction (\lstinline{sub}).

\begin{figure}[h]
\label{fig:tlc-folded}
\begin{center}
\input{tlc-folded.py}
\caption{The residual code of the TLC main loop with respect to the
\lstinline{abs} function }
\end{center}
\end{figure}

This shows a common shortcoming of partial evaluation when applied to a dynamic
language: The partial evaluator (just like an ahead-of-time compiler) cannot
make assumptions about the types of objects, which leads to poor results.
Effective dynamic compilation requires feedback of runtime information into
compile-time. This is no different for partial evaluation.

Partial evaluation (PE) comes in two flavors: \cfbolz{if we have space problems
we should kill the bits about online and offline PE}
